---
title: "Accessing Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{accessing-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Accessing data

Databrary is a data library, one specialized for storing and sharing video.
Let's see how to use `databraryr` to access data.

We'll start simply.
Let's download a test video from [volume 1 on Databrary](https://databrary.org/volume/1).

The `download_video()` function handles this for us.
Running it with the default parameters downloads a simple test video with numbers
than increment.
The file is stored in a temporary directory created by the file system using the function `tempdir()`.
The `download_asset()` function returns a character string with the full file name.

```{r, eval=FALSE}
download_video()
```

Depending on your operating system, the following commands may open the file so that you can play it with your default video player.

```{r, eval=FALSE}
nums_vid <- download_video()
system(paste0("open ", nums_vid))
```

Or, you can navigate to the temporary directory to open and play the video manually.
Use `tempdir()` to find the directory where `test.mp4` is stored.

Now, let's see what other files are shared in volume 1.
This takes a moment to run because there are *many* files in this volume.

```{r}
vol1_df <- list_volume_assets()
```

The command returns a data frame we can manipulate using standard R commands.
Here are the variables in the data frame.

```{r}
names(vol1_df)
```

The `asset_format` variable tells us the type of the data file.

```{r}
unique(vol1_df$asset_format)
```

But this isn't especially informative since the `asset_format` is a code, and we don't really know what `-800` or `6` or the other numbers refer to.
To decode it, we create a data frame of all the file formats Databrary currently recognizes.

```{r}
db_constants <- assign_constants()

formats_df <- purrr::map(db_constants$format, as.data.frame) |>
  purrr::list_rbind()

formats_df
```

From this, we see that format `-800` is an MP4-formatted video. 
There are lots of those in Volume 1.
We see that`6` is a PDF document.

Let's add a more human-readable file type to the volume 1 info.

```{r}
these_formats_df <- formats_df |>
  dplyr::select(id, extension, name) |>
  dplyr::rename(asset_type = name)

vol1_df_new <- dplyr::left_join(vol1_df, these_formats_df, by=dplyr::join_by(asset_format==id))

vol1_df_new |>
  dplyr::select(asset_id, asset_size, extension, asset_type)
```

We can summarize the number of files using the `stats::xtabs()` function:

```{r}
stats::xtabs(~ asset_type, data = vol1_df_new)
```

So, there are lots of videos and PDFs to examine.
Here is a table of the ten longest videos.

```{r}
vol1_df_new |>
  dplyr::filter(asset_type == "MPEG-4 video") |>
  dplyr::select(asset_name, asset_duration) |>
  dplyr::mutate(hrs = asset_duration/(60*60*1000)) |>
  dplyr::select(asset_name, hrs) |>
  dplyr::arrange(desc(hrs)) |>
  head(n = 10) |>
  knitr::kable(format = 'html')
```

## Accessing metadata

Imagine you are interested in knowing more about this volume, the people who created it, or the agencies that funded it.

The `list_volume_owners()` function returns a data frame with information about the people who created and "own" this particular dataset.
The function has a parameter `this_vol_id` which is an integer, unique across Databrary, that refers to the specific dataset.
The `list_volume_owners()` function uses volume 1 as the default.

```{r}
list_volume_owners()
```

The command (and many like it) can be "vectorized" using the `purrr` package.

```{r}
purrr::map(1:15, list_volume_owners) |> 
  purrr::list_rbind()
```

The `list_volume_metadata()` command gives slightly more information.

```{r}
list_volume_metadata()
```

This command can also be "vectorized."

```{r}
purrr::map(c(1:50), list_volume_metadata) |>
  purrr::list_rbind()
```

The `permission` variable indicates whether a volume is visible by others by a user.

So, if you are not logged-in to Databrary, only data that are visible to the public will be returned.
Assuming you are *not* logged-in, the above commands will show volumes with `permission` equal to 1.
The `permission` field derives from a set of constants the system uses.

```{r}
db_constants <- assign_constants()
db_constants$permission
```

The `permission` array is indexed beginning with 0.
So the 1th value is "`r db_constants$permission[2]`".
So, the `1` means that the volumes shown above are all visible to the public, and to you.

Volumes that you have not shared and are not visible to the public, will have `permission` equal to 5, or "`r db_constants$permission[6]`".
We can't demonstrate this to you because we don't have privileges on the same unshared volume, but you can try it on a volume you've created but not yet shared.

The `list_volume()` command returns even more extensive information about volume 1.
The `list_volume_funding()` command returns information about any funders listed for the project.
Again, the default volume is 1.

```{r}
list_volume_funding()
```

This can also be "vectorized."

```{r}
purrr::map(c(1:20), list_volume_funding) |>
  purrr::list_rbind()
```

The `list_volume_links()` command returns information about any external (web) links that have been added to a volume, such as to related publications or a GitHub repo.

```{r}
list_volume_links()
```
