---
title: "Accessing Databrary.org via the API"
author: "Rick O. Gilmore"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
      fig_caption: yes
vignette: >        
  %\VignetteIndexEntry{Accessing Databrary.org via the API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Vignette Info

Databrary is a powerful tool for storing and sharing video data and documentation with other researchers.
With the `databraryapi` package, it becomes even more powerful.
Rather than interact with Databrary through a web browser, users can write their own code to download participant data or even specific files.

I wrote the Databrary API so I could find out what's going on on the site, and eventually so I can streamline my own analysis and data sharing workflows.
Let's get started.

## Authenticating

Access to Databrary requires prior [registration](https://databrary.org/register) and [authorization](https://www.databrary.org/resources/agreement.html) from an institution.
The authorization process requires formal agreement by an institution.
But even before formal authorization is complete, a user can access the public materials on Databrary.
All it takes is your user name (email address) and password. Let's configure your system for access then login.

```{r authentication}
# Configure the environment for Databrary
library(databraryapi)
config_db()
login_db(stored.credentials = FALSE)
```


<!-- ## Styles -->

<!-- The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows: -->

<!--     output:  -->
<!--       rmarkdown::html_vignette: -->
<!--         css: mystyles.css -->

<!-- ## Figures -->

<!-- The figure sizes have been customised so that you can easily put two images side-by-side.  -->

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

<!-- You can enable figure captions by `fig_caption: yes` in YAML: -->

<!--     output: -->
<!--       rmarkdown::html_vignette: -->
<!--         fig_caption: yes -->

<!-- Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**. -->

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
