---
title: "Accessing Databrary.org via the API"
author: "Rick O. Gilmore"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
      fig_caption: yes
vignette: >        
  %\VignetteIndexEntry{Accessing Databrary.org via the API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(tidyverse)
```

## Vignette Info

Databrary is a powerful tool for storing and sharing video data and documentation with other researchers.
With the `databraryapi` package, it becomes even more powerful.
Rather than interact with Databrary through a web browser, users can write their own code to download participant data or even specific files.

I wrote the Databrary API so I could find out what's going on on the site, and eventually so I can streamline my own analysis and data sharing workflows.
Let's get started.

## Registering

Access to Databrary requires prior [registration](https://databrary.org/register) and [authorization](https://www.databrary.org/resources/agreement.html) from an institution.
The authorization process requires formal agreement by an institution.
But you'll create an account ID (email) and secure password when you register.
Then, when you log in with your new credentials, you'll select an existing institution (if yours is on the list), a new institution (if yours isn't), or an existing authorized investigator (if you are a student, postdoc, or collaborator) to request authorization from.

## First steps (while you await authorization)

But even before formal authorization is complete, a user can access the public materials on Databrary.
For this vignetted, we'll assume you fall into this category.

<!-- First, we need to download and install the package. -->

<!-- ```{r install} -->
<!-- install.packages("devtools") -->
<!-- library(devtools) -->
<!-- devtools::install_github("PLAY-behaviorome/databraryapi") -->
<!-- ``` -->

First, let's load the package:

```{r authentication}
library(databraryapi)
```

Then, try this command to pull data about Databrary's founders:

```{r list-people-default}
list_people()
```

Databrary assigns a unique number for each person and institution on the system called a 'party id'. 
When we run `list_people(1:50)` we are asking the system to provide us information about all of the _people_ whose party id's are between 1 and 50. Let's try it:

```{r list-people-1-50}
list_people(people.list = 1:50)
```

It's a bit slow, but you should see information about people beginning with Dylan Simon and ending with Lana Karasik.

You can also try seeing what's new on Databrary.
The `get_db_stats()` command gives you information about the newly authorized people, institutions, and newly uploaded datasets.
Try this:

```{r get-db-stats}
get_db_stats("stats")
get_db_stats("people")
get_db_stats("institutions")
get_db_stats("datasets")
```

Depending on when you run this command, there may or may not be new items.

## Once you are authorized

Congratulations!
Your institution has approved your access to Databrary's identifiable data.
Now, it's time to set-up `databraryapi` so you can access these materials.

By default, `databraryapi` stores your Databrary user account (email) and password in a secure, encrypted file that your computer's operating system uses for storing other passwords.
The `keyring` package provides this functionality.
There are alternative ways of storing user credentials, but this is the recommended one.

Once you are authorized, you will gain access to a much wider range of materials on Databrary.
When that happens, you'll load the package with `library(databraryapi)` and then

```{r login, eval=FALSE}
login_db(email = "youremail@yourinstitution.edu")
```

If everything works out, you should see a `Login successful` message at the R console.
Congratulations, you are ready to access Databrary's restricted shared information along with any private, unshared information you have access to.

## Interacting with stored data files

The purpose of a data repository is to enable researchers to store and share data.
Databrary aims to go a step further by allowing researchers to interact with -- to visualize and analyze -- the data they or others have stored on the system.
The `databraryapi` package includes several commands to make this easy to do.

Let's begin by listing all of the sessions (dates and times when data was collected) in volume 1.

```{r list-vol-1-sessions}
list_sessions()
```

So, this volume contains a lot of sessions, `r dim(list_sessions())[1]` in fact.

Let's peak inside one of them, the `list_sessions()[22, 'name']` also referenced by session ID `r list_sessions()[22,'id']`.
These session IDs are unique identifiers we'll need for accessing the files contained inside.
So, let's see what's inside session `r list_sessions()[22,'id']`.

```{r list-session-assets}
list_assets(slot = list_sessions()[22,'id'], volume = 1)
```

Notice that I took row 22 and the value of the 'id' column to specify the `slot` value for the `list_assets()` function.
I took these values from the output of `list_sessions()`.
The `id` column in this table is the *asset* or file level ID.

Clearly, there are several files here, `r dim(list_assets(slot = list_sessions()[22,'id'], volume = 1))[1]` in fact.
And all are videos.
How do I know that?
Well, the `format` column has all '-800' values in it.
That doesn't help; you say.
Oh, right, here's how to link Databrary's file or 'asset' types to their format identifiers:

```{r get-databrary-file-types}
get_supported_file_types()
```

See now that '-800' has `mimetype` of `video/mp4`?
This list shows us all of the currently supported file types.
Knowing them, we can make more meaningful searches for materials we want.

Let's find all of the 'pdf' documents in volume 1:

```{r find-pdfs-volume-1}
head(list_assets_by_type(volume = 1, type = "pdf"))
```

I used the `head()` command to show you only the top of the list because I knew there are a large number of PDF files here.

### Finding volumes with videos

Databrary is a video data library, so you may naturally be interested in what volumes have videos.
The `get_video_stats()` command comes in handy here.
Let's see about the video content available in volume 1.

```{r videos-in-volume-1}
get_video_stats(volume = 1)
```

So, there's quite a lot to view there. But what about some other volumes?
This code uses the `lapply()` command from base R to run the `get_video_stats()` command across a list of volumes.
It also uses the `rbind.fill()` command from the `plyr` package to merge the list of data frames into a single one that is easy to read.
**Note:** The following takes a few moments to run, so please be patient.

```{r videos-in-volumes-1-to-5}
plyr::rbind.fill(lapply(1:5, get_video_stats))
```

So, there's a lot of video to view and reuse.
