---
title: "databrary"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{databrary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(databraryr)
```

## Vignette Info

Databrary is a powerful tool for storing and sharing video data and documentation with other researchers.
With the `databraryr` package, it becomes even more powerful.
Rather than interact with Databrary through a web browser, users can write their own code to download participant data or even specific files.

I wrote the Databrary API so I could better understand how the site works under the hood, so that I could streamline my own analysis and data sharing workflows.
Let's get started.

## Registering

Access to most of the material on Databrary requires prior [registration](https://databrary.org/register) and [authorization](https://www.databrary.org/resources/agreement.html) from an institution.
The authorization process requires formal agreement by an institution.
But you'll create an account ID (email) and secure password when you register.
Then, when you log in with your new credentials, you'll select an existing institution (if yours is on the list), a new institution (if yours isn't), or an existing authorized investigator (if you are a student, postdoc, or collaborator) to request authorization from.

## First steps (while you await authorization)

But even before formal authorization is complete, a user can access the public materials on Databrary.
For this vignette, we'll assume you fall into this category.

First, we need to download and install the `devtools` package and then the `databraryr` package from GitHub.

```{r install}
if (!require(devtools)) {
  install.packages("devtools")
}

if (!require(databraryr)) {
  devtools::install_github("databrary/databraryr") 
}
```

It's a good idea to check that your installation worked by loading it into your local workspace.

```{r}
library(databraryr)
```

Then, try this command to pull data about Databrary's founders:

```{r list-people-default}
databraryr::list_people()
```

Note that this command returns a data frame (tibble) with columns that include the first name (`prename`), last name (`sortname`), affiliation, lab or personal website, and ORCID ID if available.

Databrary assigns a unique number for each person and institution on the system called a 'party id'. 
When we run `list_people(1:25)` we are asking the system to provide us information about all of the _people_ whose party id's are between 1 and 25. Let's try it:

```{r list-people-1-25}
databraryr::list_people(people_list = 1:25)
```

It's a bit slow, but you should see information about people beginning with Dylan Simon, the developer who designed and built most of the Databrary system, and ending with Gladys Chan, a graphic designer who created the Databrary and Datavyu logos and other graphic identity elements.

You can also try seeing what's new on Databrary.
The `get_db_stats()` command gives you information about the newly authorized people, institutions, and newly uploaded datasets.
Try this:

```{r get-db-stats}
databraryr::get_db_stats("stats")
databraryr::get_db_stats("people")
databraryr::get_db_stats("institutions")
databraryr::get_db_stats("datasets")
```

Depending on when you run this command, there may or may not be new items.

## Once you are authorized

Congratulations!
Your institution has approved your access to Databrary's identifiable data.
Now, it's time to set-up `databraryr` so you can access these materials.

By default, `databraryr` stores your Databrary user account (email) and password in a secure, encrypted file that your computer's operating system uses for storing other passwords.
The `keyring` package provides this functionality.
There are alternative ways of storing user credentials, but this is the recommended one.

Once you are authorized, you will gain access to a much wider range of materials on Databrary.
When that happens, you'll load the package with `library(databraryr)` and then run `login_db(email = "<YOUR_EMAIL@PROVIDER.COM>")`, substituting your actual Databrary account for `<YOUR_EMAIL@PROVIDER.COM>`, of course.
I prefer to give the package name when I do this, so the following is how I do the same thing.

```{r login, eval=FALSE}
databraryr::login_db(email = "youremail@yourinstitution.edu")
```

If everything works out, you should see a `Login successful` message at the R console.
Congratulations, you are ready to access Databrary's restricted shared information along with any private, unshared information you have access to.

**NOTE**: You can save yourself some time if you store your Databrary login (email) as an
environment variable:

1. Install the `usethis` package via `install.packages('usethis')`.
2. Run `usethis::edit_r_environ()`. This will open your `.Renviron` file in a new window.
3. Edit the `.Renviron` file by adding a line with `DATABRARY_LOGIN="youremail@yourinstitution.edu"`, substituting your actual Databrary login email.
4. Save the file, and restart R.

Now, you can run `Sys.getenv("DATABRARY_LOGIN")` and it will return your Databrary login.

And going forward, you can use `Sys.getenv("DATABRARY_LOGIN")` wherever you would enter your Databrary login:

```{r}
databraryr::login_db(email = Sys.getenv("DATABRARY_LOGIN"))
```

The package also has a log out command.

```{r}
databraryr::logout_db()
```

