---
title: "databrary"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{databrary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(databraryr)
```

## Vignette Info

Databrary is a powerful tool for storing and sharing video data and documentation with other researchers.
With the `databraryr` package, it becomes even more powerful.
Rather than interact with Databrary through a web browser, users can write their own code to download participant data or even specific files.

I wrote the Databrary API so I could better understand how the site works under the hood, so that I could streamline my own analysis and data sharing workflows.
Let's get started.

## Registering

Access to most of the material on Databrary requires prior [registration](https://databrary.org/register) and [authorization](https://www.databrary.org/resources/agreement.html) from an institution.
The authorization process requires formal agreement by an institution.
But you'll create an account ID (email) and secure password when you register.
Then, when you log in with your new credentials, you'll select an existing institution (if yours is on the list), a new institution (if yours isn't), or an existing authorized investigator (if you are a student, postdoc, or collaborator) to request authorization from.

## First steps (while you await authorization)

But even before formal authorization is complete, a user can access the public materials on Databrary.
For this vignette, we'll assume you fall into this category.

First, we need to download and install the `devtools` package and then the `databraryr` package from GitHub.

```{r install}
if (!require(devtools)) {
  install.packages("devtools")
}

if (!require(databraryr)) {
  devtools::install_github("databrary/databraryr") 
}
```

It's a good idea to check that your installation worked by loading it into your local workspace.

```{r}
library(databraryr)
```

Then, try this command to pull data about Databrary's founders:

```{r list-people-default}
databraryr::list_people()
```

Note that this command returns a data frame (tibble) with columns that include the first name (`prename`), last name (`sortname`), affiliation, lab or personal website, and ORCID ID if available.

Databrary assigns a unique number for each person and institution on the system called a 'party id'. 
When we run `list_people(1:50)` we are asking the system to provide us information about all of the _people_ whose party id's are between 1 and 50. Let's try it:

```{r list-people-1-50}
databraryr::list_people(people_list = 1:50) # databraryr::list_people(1:50) would also work
```