---
title: "Accessing Databrary.org via the API"
author: "Rick O. Gilmore"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
      fig_caption: yes
vignette: >        
  %\VignetteIndexEntry{Accessing Databrary.org via the API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(tidyverse)
```

## Vignette Info

Databrary is a powerful tool for storing and sharing video data and documentation with other researchers.
With the `databraryapi` package, it becomes even more powerful.
Rather than interact with Databrary through a web browser, users can write their own code to download participant data or even specific files.

I wrote the Databrary API so I could find out what's going on on the site, and eventually so I can streamline my own analysis and data sharing workflows.
Let's get started.

## Registering

Access to Databrary requires prior [registration](https://databrary.org/register) and [authorization](https://www.databrary.org/resources/agreement.html) from an institution.
The authorization process requires formal agreement by an institution.
But you'll create an account ID (email) and secure password when you register.
Then, when you log in with your new credentials, you'll select an existing institution (if yours is on the list), a new institution (if yours isn't), or an existing authorized investigator (if you are a student, postdoc, or collaborator) to request authorization from.

## First steps (while you await authorization)

But even before formal authorization is complete, a user can access the public materials on Databrary.
For this vignetted, we'll assume you fall into this category.

<!-- First, we need to download and install the package. -->

<!-- ```{r install} -->
<!-- install.packages("devtools") -->
<!-- library(devtools) -->
<!-- devtools::install_github("PLAY-behaviorome/databraryapi") -->
<!-- ``` -->

First, let's load the package:

```{r load-databraryapi}
library(databraryapi)
```

Then, try this command to pull data about Databrary's founders:

```{r list-people-default}
list_people()
```

Databrary assigns a unique number for each person and institution on the system called a 'party id'. 
When we run `list_people(1:50)` we are asking the system to provide us information about all of the _people_ whose party id's are between 1 and 50. Let's try it:

```{r list-people-1-50}
list_people(people.list = 1:50)
```

It's a bit slow, but you should see information about people beginning with Dylan Simon and ending with Lana Karasik.

You can also try seeing what's new on Databrary.
The `get_db_stats()` command gives you information about the newly authorized people, institutions, and newly uploaded datasets.
Try this:

```{r get-db-stats}
get_db_stats("stats")
get_db_stats("people")
get_db_stats("institutions")
get_db_stats("datasets")
```

Depending on when you run this command, there may or may not be new items.

## Once you are authorized

Congratulations!
Your institution has approved your access to Databrary's identifiable data.
Now, it's time to set-up `databraryapi` so you can access these materials.

By default, `databraryapi` stores your Databrary user account (email) and password in a secure, encrypted file that your computer's operating system uses for storing other passwords.
The `keyring` package provides this functionality.
There are alternative ways of storing user credentials, but this is the recommended one.

Once you are authorized, you will gain access to a much wider range of materials on Databrary.
When that happens, you'll load the package with `library(databraryapi)` and then

```{r login, eval=FALSE}
login_db(email = "youremail@yourinstitution.edu")
```

If everything works out, you should see a `Login successful` message at the R console.
Congratulations, you are ready to access Databrary's restricted shared information along with any private, unshared information you have access to.
